Plant-pollintor networks worldwide are composed by the same specific building

### 2021-08-05 Draft/Draft.Rmd

Finally, we studied which motif combinations of functional groups (up to five nodes) are over or under-represented in 95% of our empirical networks (57 out of 60 arrangements, due to computational limitations to identify all the nodes in the motifs of the three networks with the highest number of links). To do so, for each of the 53,250 possible motif combinations, we estimated the observed and the expected probability of finding that combination in empirical networks, respectively. Then, we determined whether the observed probabilities are likely to come from the expected probabilities or not. To calculate the observed probability of a given combination $i$, $p_i^O$, we divided the number of times that $i$ appears in our set of empirical networks ($n_i^O$) by the sum of the number of times that each possible combination appeared: $p_i^O = n_i^O / \sum_{k=1}^{53,250} n_k^O$, and, consequently, $\sum_{k=1}^{53,250} p_k^O = 1$. To estimate the expected probability of a given motif combination $i$, $p_i^E$, firstly, we calculated the probability of finding a given functional group $x$ in the position $\alpha$ of $i$, $p_i(x,\alpha)$. Then, by assuming the independency of $p_i(x,\alpha)$, we computed the expected probability of the combination $i$ as the product of the probability of its pairs $(x,\alpha)$, that is, $p_i^E = \prod _{(x,\alpha )\in i} p_i(x,\alpha)$. To obtain $p_i(x,\alpha)$, we proceeded as follows. First, we used the _bmotif_ package [@bmotifsimmons2019] to calculate the number of times (or absolute frequency) that the functional group $x$ appears in the position $\alpha$ in each empirical network $\eta$, denoted as $n_i^{\eta}(x,\alpha)$. Then, by controlling the variation at the network level, we assessed the average value of the absolute frequency of functional group $x$ appears in the position $\alpha$, $n_i(x,\alpha) = E\left [ n_i^{\eta}(x,\alpha) \right ]$, and estimated $p_i(x,\alpha)$ as $p_i(x,\alpha) = n_i(x,\alpha)/\sum_{k}n_i(k,\alpha)$. To obtain $n_i(x,\alpha)$, we fitted a LMM per motif position, where the response variable was the number of times that a given position was observed per network, the explanatory variable was functional group identifier, and the random intercept was given by network identifiers nested within the study identifiers. We used the ‘lmer’ package [@bates2015] to fit the LMMs models in our analyses.

Once we obtained $p_i^O$ and $p_i^E$, we used simulation to determine whether the former is likely to come from the latter or not, since the large number of possible motif combinations and the small probabilities for some of them advise against using an Exact test of goodness-of-fit or a Chi-square one. Specifically, we created 1,000 random samples with repetition of possible motif combinations, where each sample contained 10 million elements and, for each combination, the probability of being selected was equal to its expected probability. From those random samples, we extracted the mean and the standard deviation of the expected probability of $i$, $E\left [ p_i^E \right ]$ and $\sigma\left [ p_i^E \right ]$, respectively, and calculated the z-scores of $p_i^O$ as $z_i^O = \left ( p_i^O - E\left [ p_i^E \right ] \right )/\sigma \left [ p_i^E \right ]$, for those motif combinations with $p_i^O > 0$. Following the usual interpretation for z-scores, combinations with $z_i^O > 1.96$ are over-represented, whereas those with $z_i^O < -1.96$ are under-represented, at the 95%  confidence level. Notice that we focused on combinations with $p_i^O > 10^{-7}$ because, beyond the interest that can arouse the combinations that appear at most once, we do not have enough numerical resolution to accurately detect whether or not those combinations are under-represented (due to the limited size of our random samples).

>>>>

Finally, we studied which motif combinations of functional groups (up to five nodes) are over or under-represented in 95% of our empirical networks (57 out of 60 arrangements, due to computational limitations to identify all the nodes in the motifs of the three networks with the highest number of links). To do so, for each of the 53,250 possible motif combinations, we estimated the observed and the expected probability of finding that combination in empirical networks, respectively. Then, we determined whether the observed probabilities are likely to come from the expected probabilities or not. To calculate the observed probability of a given combination $i$, $p_i^O$, we divided the number of times that $i$ appears in our set of empirical networks ($n_i^O$) by the sum of the number of times that each possible combination appeared: $p_i^O = n_i^O / \sum_{k=1}^{53,250} n_k^O$, and, consequently, $\sum_{k=1}^{53,250} p_k^O = 1$. To estimate the expected probability of a given motif combination $i$, $p_i^E$, firstly, we calculated the probability of finding a given functional group $x$ in the position $\alpha$ of $i$, $p_i(x,\alpha)$. Then, by assuming the independency of $p_i(x,\alpha)$, we computed the expected probability of the combination $i$ as the product of the probability of its pairs $(x,\alpha)$, that is, $p_i^E = \prod _{(x,\alpha )\in i} p_i(x,\alpha)$. To obtain $p_i(x,\alpha)$, we proceeded as follows. First, we calculated the number of times that the functional group $x$ appears in the position $\alpha$ in each empirical network $\eta$, denoted as $n_i^{\eta}(x,\alpha)$. Then, by controlling the variation at the network level, we assessed the average value of the absolute frequency of functional group $x$ appears in the position $\alpha$, $n_i(x,\alpha) = E\left [ n_i^{\eta}(x,\alpha) \right ]$, and estimated $p_i(x,\alpha)$ as $p_i(x,\alpha) = n_i(x,\alpha)/\sum_{k}n_i(k,\alpha)$. To obtain $n_i(x,\alpha)$, we fitted a LMM per motif position, where the response variable was the number of times that a given position was observed per network, the explanatory variable was functional group identifier, and the random intercept was given by network identifiers nested within the study identifiers. We used the ‘lmer’ package [@bates2015] to fit the LMMs models in our analyses.

Once we obtained $p_i^O$ and $p_i^E$, we used simulation to determine whether the former is likely to come from the latter or not, since the large number of possible motif combinations and the small probabilities for some of them advise against using an Exact test of goodness-of-fit or a Chi-square one. Specifically, we created 1,000 random samples with repetition of possible motif combinations, where each sample contained 10 million elements and, for each combination, the probability of being selected was equal to its expected probability. From those random samples, we extracted the mean and the standard deviation of the expected probability of $i$, denoted as $E\left [ p_i^E \right ]$ and $\sigma\left [ p_i^E \right ]$, respectively, and calculated the z-scores of $p_i^O$ as $z_i^O = \left ( p_i^O - E\left [ p_i^E \right ] \right )/\sigma \left [ p_i^E \right ]$, for those motif combinations with $p_i^O > 0$. According to the usual interpretation of z-scores, combinations with $z_i^O > 1.96$ are over-represented, whereas those with $z_i^O < -1.96$ are under-represented, at the 95%  confidence level. Notice that we focused on combinations with $p_i^O > 10^{-7}$ because, beyond the interest that can arouse the combinations that appear at most once, we do not have enough numerical resolution to accurately detect whether or not those combinations are under-represented (due to the limited size of our random samples).
